\textls[-0]{\thetitle}

### 2023-02-08 pldi2023/review-reply-densed.md

  However, our translation doesn't translate this unconventional Wtype
  formulation back to the conventional one. So the target calculus after
  translation is **MLTT + this unconventional Wtype**. In that case, translation
  can only function as a guide to the plugin implementation, and
  **pedantically** is not enough to show the consistency/canonicitiy for FMLTT
  because **pedantically** target calculus is not proven to be
  consistency/canonicity.

  We agree with Reviewer B's insight -- we will get consistency/canonicitiy when
  we translate unconventional Wtype into the conventional one. The reason we
  didn't choose to do so because we expect this translation a lot more verbose
  than the current proof because of the simplicity of Wtype itself compared to
  the rich functionality provided by (fake-)Agda's Inductive Facility. We only
  use the latter when constructing consistency/canonicity model.

### Other questions

* (Reviewer A/B) Why is W(t) a term and not a type? What are bold-W and bold-P needed for?

@@

  𝕎(t) is a typo and should be deleted. Thanks for pointing out.

  ℙ is another technical detail of the system. Its functionality is to transform
  a linkage(overridable/extensible) into a module(sigma type). The reason it is
  irreplacable is because, proper abstraction cannot happen on linkage (𝕃) but
  only on a module (ℙ). For example, Line 744 shows we can prove `ℙ(σ₅) ⊢ s₆ :
  A₆[p¹]` but generally `𝕃(σ₅) ⊢ s₆ : A₆[p¹]` not provable for non-trivial
  `A₆`.

* *(Reviewer B) How will our plugin react when trying to mixin the contradictory features? For example, since STLC with eith polymorphism or references enjoy type soundness, but their composition doesn't.*

<!-- I think we still need to clarify we doesn't support extending these -->
We currently doesn't support extend STLC with polymorphism and reference, which requires extending existent inductive family with new indices.

Hypothetically speaking, We can expect our plugin will generate unprovable proof obligation under mixin, hindering qed'ed the proposition and thus closing the family.

@@

Another exmaple would be extend *a family of STLC and its termination proof* with general recursion feature. Our plugin will generate unprovable proof obligation inside the reducibility argument for the fixpoint feature.

* (Reviewer D) **"What goes wrong if a user tries to define a term with the `tm_rect` type using standard pattern matching? Is it rejected? In the latter case, how does the plugin prevent other extensible definitions from referring to it?"**

@@

  Coq will reject on the pattern matching on `tm`. For example, Figure 4, Line
  494 `Module Type STLC°tm` shows the abstraction/wrapped interface around the
  given inductive `tm`. The following field definition will be type-checked
  based on this interface, and for Coq it will only consider `tm` as an
  arbitrary type and will fail to pattern match it.

* (Reviewer D) **""In the event that fpop cannot infer where the programmer intends to place a new field, annotation is required." I did not understand what it means to 'place' a field-- when does this occur,  and what do these annotations look like?"**

For example
```C
Family A {Field a; Field b; Field c;}

Family B extends A {
  (* Inherit a. *) 
  (* this vernacular command line will make a difference once uncommented. Before uncommented it, a2 will be the first field of the Family B.
      The fields are order sensitive, especially when the definition of a2 dependent on a, placing a2 before a will cause program rejected. 
      This ``annotation`` require the programmers effort. *) 
  Field a2.
}
```

@@

<!-- I am not sure if I want to mention overridable/pins  -->

* (Reviewer D) **"First, a module named STLC◦subst◦Cases is generated interactively: every time the programmer completes a..." This is also confusing-- why is the programmer involved in the translation? Is it that modules are being generated in the background, while commands are being processed?**

We should be more clear that we want to emphasize, whenever one vernacular command is emitted, our plugin will translate and type-check. Basically type-checking happens together with interactive theorem proving, as opposed to non-interactively -- where the type-checking happens after a whole family is closed

Yes, modules/functors are generated in the background after each command is emitted.

### 2023-02-08 pldi2023/review-reply-densed.md

* "Seems related to type theories for modules, but doesn't discuss the connection"

Finally, I note that a family has similarities with an ML-style module,
including the fact that some of its members are opaque while others are
abstract. This is modelled through the use of singleton types, just as
dependent type theories for modules do (e.g., Dreyer, Crary, Harper, POPL
2003), which are also extensions of MLTT. Furthermore, there has been work on
extending ML modules with mixin composition (Dreyer, Rossberg, ICFP 2008).
ML-style modules are the basis of Coq's module system as well, so I would have
expected some comparison with this line of work.

  > We need to say something about "A type system for higher-order modules"

Generally speaking, compared to ML-style module, Family focuses on overridability and code inheritance. The latter can be modeled by the functor in a verbose way. Module also has a clear distinction on the implementation and its signature, while family doesn't -- a given family is usually fixed with one signature, closer to mundane classes.

Compared to the paper [DCH 2003], their problem formulation is more mature and they gear towards real-life programming experience. For example, they aim at generativity (a feature for nominality and side effect); subtyping (happens during signature matching); and phase distinction (for compilation). Our current paper handles family in a structural style; doesn't relate different signatures at all; and we work in a full-dependent type setting where mixing static and dynamic phase is acceptable.

In fact, for plugin development, the family is compiled into module/functors. In meta-theory, we use sigma type as a conceptually simpler representation of modules. We also use singleton type to expose concrete type information in a family inspired by their work.

### 2023-02-08 pldi2023/review-reply-densed.md

As Reviewer A/D points out, to inductiely eliminate data we currently only are equipped with `FInduction` at top level. We only have `fdiscriminate` and `finjection` for deducing intermediate fact. The key problem is the lack of certain nested pattern matching.

<!-- The idea is 
    fill in the new hole once pattern matching needs to extend the clauses
    lifting nested induction to top level is not necessary
 -->

We acknowledge this limitation of our current research and are trying to advance more in this direction. We hypothesize that code involving nested induction/pattern matching an be reused by fill in the hole after extension upon inductive type. The plugin will generate extra proof obligation to fill this hole. This certainly requires significant amount of engineering effort.

* (Reviewer A) Some comparison between our work and [DCH 2003] [DR 2008]?

Generally speaking, compared to ML-style module, Family focuses on overridability and code inheritance. The latter can be modeled by the functor in a verbose way. Module also has a clear distinction between the implementation and its signature, while family doesn't -- a given family is usually fixed with one signature, closer to OO classes.

### 2023-02-08 pldi2023/review-reply-densed.md

We thank all the reviewers for their critical and encouraging feedback.
Below we address the main concerns, which might be paraphrased for the space.

* (Review A) __"In terms of ergonomics, I wonder how powerful and usable" this is ... ; (Reviewer D) "Another place where the current approach seems to be lacking is the restriction of induction / case analysis to the top level."__

  As Reviewer A/D highlights, the current method for inductively eliminating data only involves using "FInduction" at the top level. We have "fdiscriminate" and "finjection" for deducing intermediate facts, but the key problem is the absence of proper nested pattern matching.

@@

  <!-- The idea is 
      fill in the new hole once pattern matching needs to extend the clauses
      lifting nested induction to top level is not necessary
  -->

  We acknowledge this limitation of our current research and are trying to advance more in this direction. We hypothesize that code involving nested induction/pattern matching can be reused by fill in the hole after extension happens upon inductive type. The plugin will generate extra proof obligation to fill this hole. This certainly requires significant amount of engineering effort.

* (Review A) __Some comparison between this paper and [DCH 2003] [DR 2008]?__

@@

  Generally speaking, compared to ML-style module, Family focuses on overridability and code inheritance. The latter can be modeled by the functor in a verbose way. Module also has a clear distinction between the implementation and its signature, while family doesn't -- a given family is usually fixed with one signature, closer to OO classes.

  Compared to the paper [DCH 2003], their problem formulation is more mature and they gear towards real-life programming experience. For example, they aim at generativity (a feature for nominality and side effect); subtyping (happens during signature matching); and phase distinction (for compilation). Our current paper handles family in a structural style; doesn't relate different signatures at all; and we work in a full-dependent type setting where mixing static and dynamic phase is acceptable.

  In fact, for plugin development, the family is compiled into module/functors. In meta-theory, we use sigma type as a conceptually simpler representation of modules. We also use singleton type to expose concrete type information in a family inspired by their work. So our work is heavily influenced by ML-modules.

* (Review A/B/C/D) __Section 5 FMLTT is too dense, "incrediably terse" and inappropriately presented to give the audience enough intuition. "More worrisome, the full proof is *only* given in a notation heavy, Agda-style format"__

@@

  Thank you for your feedback. 
  We sympathize with this impression on Section 5, and we agree it is dense, hard to read and not giving too much inspiration for the readers.

  We plan to expand it in the main text to clarify the questions the reviewers have, and also use appendices to explain the formalization and the proof in greater details aiming for accessibility.

* (Review B) __The `A` of _TyEq/PK/Add_
  and _LSig/Add_ where `A` decides what to make available to later definitions. Is `A` non-determinstic? How does plugin decides `A`?__

@@

  The typing rules do not commit to a particular choice of $A$, as the reviewer
  points out.
  In practice, this A is decided by implementation. Our plugin always choose a "default" one which is just making all inductive type into an opaque type (Line 791) and make sure other parts stay the "same".

  Taking (Line 745) Figure. 8 as an example, `σ₅` has `tm : 𝕊(W(τₜₘ))` and `A`
  has `tm : 𝕌` instead. `s` will make sure other fields stay the same type. To
  show this explicitly at the plugin level, we look at Figure 4 (Line 494)
  `Module Type STLC°tm` (corresponding to `A`), where `tm : Set` (corresponding
  to `tm : 𝕌`). With this interface `STLC°tm` we cannot pattern match any term
  of type `tm : 𝕌` thus doing abstraction successfully.

  Generally speaking, all the (extensible) inductive type will be simply
  "wrapped" by a module type only exposing the
  constructor (with no eliminators), just like how we generate `STLC°tm`.

* (Review B/D) __"I find it somewhat surprising that neither the FPOP plugin / implementation itself, nor the code of the case studies, is provided."__

@@

  We will certainly make the FPOP plugin/implementation available.

* (Review B) __"I think the paper would also be improved by delineating the limitations of the current implementation / metatheory.."__; (Review C) __"Be upfront about this limitation ..."__; (Review D)__ "The biggest of these (which is not mentioned in the paper), is that the presence of open terms complicates typechecking dependently typed terms."__; (Review D) __"It would be nice to add an explicit comment about the trusted code bases of any developments"__

  Thank you for your feedback. We will clearly state the limitations of the
  current implementation and meta-theory in the final revision.

  In particular, we will emphasize the issue raised by Review D regarding equality coercion
  and the pervasive propositional equality on extensible inductive families.
  This can result in unnecessary equality coercion when `denoteTy t` is not
  definitionally equal to `nat`.

@@

  We believe that this opens up new opportunities for future work in both theory
  and practice, (a). such as a normalization algorithm for an extended calculus
  that allows for more convertibility upon elimination of the extensible
  inductive type; (b). and a proper proof assistant that supports native
  extensible inductive types instead of the current encoding via our plugin,
  thus allowing for proper reduction of open terms. This will change the kernel
  of proof assistant for sure.

  In the final revision, we will also clarify that our plugin only translates
  code into vanilla Coq terms without expanding the trusted codebase.

* (Review C) __The utility of the FMLTT formalization seems unclear since consistency and canonicity can already be deduced by the translation to MLTT following the approach in FPOP__

@@

  The technical detail here is that, FMLTT also has this unconventional Wtype
  formulation. While each classic Wtype is defined by a pair of type `A ⊢ B`,
  our unconventional Wtype is defined by a list of pair of types `Aᵢ ⊢ Bᵢ`, each
  pair corresponds to one constructor.

  One way to interpret it is the classical formluation encoding this list using
  a big sum type. So the conventional and unconventional Wtype can be translated
  to each other.

  However, our translation doesn't translate this unconventional Wtype
  formulation back to the conventional one. So the target calculus after
  translation is **MLTT + this unconventional Wtype**. In that case, translation
  can only function as a guide to the plugin implementation, and
  **pedantically** is not enough to show the consistency/canonicitiy for FMLTT
  because, **pedantically**, target calculus is not proven to be
  consistency/canonicity.

  We agree with Reviewer B's insight -- we will get consistency/canonicitiy when
  we translate unconventional Wtype into the conventional one. The reason we
  didn't choose to do so is that we expect this translation a lot more verbose
  than the current proof because of the simplicity of Wtype itself compared to
  the rich functionality provided by (fake-)Agda's Inductive Facility. We only
  use the latter when constructing consistency/canonicity model.

* (Review A/B) __In the conclusion of rule Tyeq/Casety, R should be T?__

Yes. Thank you for noticing this typo.

* (Review A) __"Why is W(t) a term and not a type? What are bold-W and bold-P needed for?"__

@@

  In dependent type system, a type is also a term. i.e. we have to allow `Γ ⊢ W(t) : 𝕌`.

  𝕎(t) is a typo and should be deleted. Thanks for pointing out.

  ℙ is another technical detail of the system. Its functionality is to transform
  a linkage(overridable/extensible) into a module(sigma type). The reason it is
  irreplacable is because, proper abstraction cannot happen on linkage (𝕃) but
  only on a module (ℙ). For example, Line 744 shows we can prove `ℙ(σ₅) ⊢ s₆ :
  A₆[p¹]` but generally `𝕃(σ₅) ⊢ s₆ : A₆[p¹]` not provable for non-trivial
  `A₆`.

* *(Review B) How will our plugin react when trying to mixin the contradictory features? For example, since STLC with eith polymorphism or references enjoy type soundness, but their composition doesn't.*

@@

  <!-- I think we still need to clarify we doesn't support extending these -->
  We currently doesn't support extend STLC with polymorphism and reference, which requires extending existent inductive family with new indices.

  Hypothetically speaking, We can expect our plugin will generate unprovable proof obligation under mixin, hindering qed'ed the proposition and thus closing the family.

  Another exmaple would be extend *a family of STLC and its termination proof* with general recursion feature. Our plugin will generate unprovable proof obligation inside the reducibility argument for the fixpoint feature.

* (Review D) __"What goes wrong if a user tries to define a term with the `tm_rect` type using standard pattern matching? Is it rejected? In the latter case, how does the plugin prevent other extensible definitions from referring to it?"__

@@

  Coq will reject on the pattern matching on `tm`. For example, Figure 4, Line 494 `Module Type STLC°tm` shows the abstraction/wrapped interface around the given inductive `tm`. The following field definition will be type-checked
  based on this interface, and for Coq it will only consider `tm` as an
  arbitrary type and will fail to pattern match it.

* (Review D) __""In the event that fpop cannot infer where the programmer intends to place a new field, annotation is required." I did not understand what it means to 'place' a field-- when does this occur,  and what do these annotations look like?"__

@@

  For example
```C
Family A {Field a; Field b; Field c;}

Family B extends A {
  (* Inherit a. *) 
  (* this vernacular command line will make a difference once uncommented. Before uncommented it, a2 will be the first field of the Family B.
      The fields are order sensitive, especially when the definition of a2 dependent on a, placing a2 before a will cause program rejected. 
      This ``annotation`` require the programmers effort. *) 
  Field a2.
}
```

<!-- I am not sure if I want to mention overridable/pins  -->

* (Review D) __"First, a module named STLC◦subst◦Cases is generated interactively: every time the programmer completes a..." This is also confusing-- why is the programmer involved in the translation? Is it that modules are being generated in the background, while commands are being processed?__

### 2023-02-08 pldi2023/review-reply-densed.md

* (Review B/D) __"I find it somewhat surprising that neither the FPOP plugin / implementation itself, nor the code of the case studies, is provided."__

  We will certainly make the FPOP plugin/implementation available.

* (Review B) __"I think the paper would also be improved by delineating the limitations of the current implementation / metatheory.."__; (Review C) __"Be upfront about this limitation ..."__; (Review D)__ "The biggest of these (which is not mentioned in the paper), is that the presence of open terms complicates typechecking dependently typed terms."__; (Review D) __"It would be nice to add an explicit comment about the trusted code bases of any developments"__

  Thank you for your feedback. We will clearly state the limitations of the
  current implementation and meta-theory in the final revision. 
  <!-- do we want to mention overridability/pins? -->
  In particular, we will emphasize the issue raised by Review D regarding equality coercion
  and the pervasive propositional equality on extensible inductive families.
  This can result in unnecessary equality coercion when `denoteTy t` is not
  definitionally equal to `nat`.

@@

  In dependent type system, a type is also a term. i.e. we have to allow `Γ ⊢ W(t) : 𝕌`.

  𝕎(t) is a typo and should be deleted. Thanks for pointing out.

  ℙ is another technical detail of the system. Its functionality is to transform
  a linkage(overridable/extensible) into a module(sigma type). The reason it is
  irreplacable is because, proper abstraction cannot happen on linkage (𝕃) but
  only on a module (ℙ). For example, Line 744 shows we can prove `ℙ(σ₅) ⊢ s₆ :
  A₆[p¹]` but generally `𝕃(σ₅) ⊢ s₆ : A₆[p¹]` not provable for non-trivial
  `A₆`.

* (Review B) __How will our plugin react when trying to mixin the contradictory features? For example, since STLC with eith polymorphism or references enjoy type soundness, but their composition doesn't.__

@@

  For example
```C
Family A {Field a; Field b; Field c;}

Family B extends A {
  (* Inherit a. *) 
  (* this vernacular command line will make a difference once uncommented. Before uncommented it, a2 will be the first field of the Family B.
      The fields are order sensitive, especially when the definition of a2 dependent on a, placing a2 before a will cause program rejected. 
      This ``annotation`` require the programmers effort. *) 
  Field a2.
}
```

* (Review D) __"First, a module named STLC◦subst◦Cases is generated interactively: every time the programmer completes a..." This is also confusing-- why is the programmer involved in the translation? Is it that modules are being generated in the background, while commands are being processed?__

  We should be clearer that, the point we want to emphasize is, whenever one vernacular command is emitted, our plugin will translate and type-check. Basically type-checking happens together with interactive theorem proving, as opposed to non-interactively -- where the type-checking only happens after a whole family is closed.

### 2023-02-08 pldi2023/review-reply-densed.md

We thank all the reviewers for their critical and encouraging feedback.
Below we address the main concerns, which might be paraphrased for the space.

* (Review A) __"In terms of ergonomics, I wonder how powerful and usable" this is ... ; (Reviewer D) "Another place where the current approach seems to be lacking is the restriction of induction / case analysis to the top level."__

  As Reviewer A/D highlights, the current method for inductively eliminating data only involves using "FInduction" at the top level. We have "fdiscriminate" and "finjection" for deducing intermediate facts, but the key problem is the absence of proper nested pattern matching.

  <!-- The idea is 
      fill in the new hole once pattern matching needs to extend the clauses

      lifting nested induction to top level seems not necessary
  -->

@@

* (Review A/B/C/D) __Section 5 FMLTT is too dense, "incrediably terse" and inappropriately presented to give the audience enough intuition. "More worrisome, the full proof is *only* given in a notation heavy, Agda-style format"__

  Thank you for your feedback. 
  We sympathize with this impression on Section 5, and we agree it is dense, hard to read and not giving too much inspiration for the readers.

  We plan to expand it in the main text to clarify the questions the reviewers have, and also use appendices to explain the formalization and the proof in greater details aiming for accessibility.

* (Review-B) __Rules TYEQ/PK/ADD and LSIG/ADD do not seem to prescribe how to choose $A$.__

  The FMLTT typing rules do not commit to a particular choice of the context type $A$,
  though there exists a simple algorithm for picking the right $A$ (lines 785–787): a
  field is checked in a context of type $A$ that hides W-type signatures behind
  $\mathbb{U}$, unless that field invokes a constructor or eliminator of some
  W-type. Intuitively, the type system is sound without this algorithm because
  it restricts the ways a field can be used if a wrong $A$ is chosen.

<!--
   In practice, this A is decided by implementation. Our plugin always choose a "default" one which is just making all inductive type into an opaque type (Line 791) and make sure other parts stay the "same".

@@

  Taking (Line 745) Figure. 8 as an example, `σ₅` has `tm : 𝕊(W(τₜₘ))` and `A`
  has `tm : 𝕌` instead. `s` will make sure other fields stay the same type. To
  show this explicitly at the plugin level, we look at Figure 4 (Line 494)
  `Module Type STLC°tm` (corresponding to `A`), where `tm : Set` (corresponding
  to `tm : 𝕌`). With this interface `STLC°tm` we cannot pattern match any term
  of type `tm : 𝕌` thus doing abstraction successfully.

  Generally speaking, all the (extensible) inductive type will be simply
  "wrapped" by a module type only exposing the
  constructor (with no eliminators), just like how we generate `STLC°tm`.
-->

* (Reviews-B&D) __The plugin implementation is not included as a supplement.__

  We will ensure the release of our research artifacts for public access.

### 2023-02-08 pldi2023/review-reply-densed.md

  Generally speaking, compared to ML-style module, Family focuses on overridability and code inheritance. The latter can be modeled by the functor in a verbose way. Module also has a clear distinction between the implementation and its signature, while family doesn't -- a given family is usually fixed with one signature, closer to OO classes.

  Compared to the paper [DCH 2003], their problem formulation is more mature and they gear towards real-life programming experience. For example, they aim at generativity (a feature for nominality and side effect); subtyping (happens during signature matching); and phase distinction (for compilation). Our current paper handles family in a structural style; doesn't relate different signatures at all; and we work in a full-dependent type setting where mixing static and dynamic phase is acceptable.

  In fact, for plugin development, the family is compiled into module/functors. In meta-theory, we use sigma type as a conceptually simpler representation of modules. We also use singleton type to expose concrete type information in a family inspired by their work. So our work is heavily influenced by ML-modules.

  Our module and mixin has the similar semantic of that from [DCH 2003] [DR 2008]. However, compared to their work, we mainly focus on the perspect of the extensible inductive type and (exhaustiveness checking of) the corresponding recursors. Even in the case of mixin, we consider the consequent mixin of the inductive type and recursors.

### 2023-02-08 pldi2023/review-reply-densed.md

  We will ensure the release of our research artifacts for public access.

* (Review B) __"I think the paper would also be improved by delineating the limitations of the current implementation / metatheory.."__; (Review C) __"Be upfront about this limitation ..."__; (Review D)__ "The biggest of these (which is not mentioned in the paper), is that the presence of open terms complicates typechecking dependently typed terms."__; (Review D) __"It would be nice to add an explicit comment about the trusted code bases of any developments"__

  Thank you for your feedback. We will clearly state the limitations of the
  current implementation and meta-theory in the final revision. 
  <!-- do we want to mention overridability/pins? -->
  In particular, we will emphasize the issue raised by Review D regarding equality coercion
  and the pervasive propositional equality on extensible inductive families.
  This can result in unnecessary equality coercion when `denoteTy t` is not
  definitionally equal to `nat`.

  We believe that this opens up new opportunities for future work in both theory
  and practice, (a). such as a normalization algorithm for an extended calculus
  that admits more conversion equation upon elimination of the extensible
  inductive type; (b). and a proper proof assistant that supports native
  extensible inductive types instead of the current encoding via our plugin,
  thus allowing for proper reduction of open terms. This will change the kernel
  of proof assistant for sure.

### 2023-02-08 pldi2023/review-reply-densed.md

We thank all the reviewers for their critical and encouraging feedback.
Below we address the main concerns, which might be paraphrased for the space.

* (Review A) __"In terms of ergonomics, I wonder how powerful and usable" this is ... ; (Reviewer D) "Another place where the current approach seems to be lacking is the restriction of induction / case analysis to the top level."__

  As Reviewer A/D highlighted, the current method for inductively eliminating data only involves using "FInduction" at the top level. We have "fdiscriminate" and "finjection" for deducing intermediate facts, but the key problem is the absence of proper nested pattern matching.

@@

  <!-- The idea is 
      fill in the new hole once pattern matching needs to extend the clauses

      lifting nested induction to top level seems not necessary
  -->

  We acknowledge this limitation of our current research and are trying to advance more in this direction. We hypothesize that we can reuse code involving nested induction/pattern matching, by filling in the new matching clauses after extension happens upon the inductive type. The plugin will generate an extra proof obligation to fill this hole. It certainly requires a significant amount of engineering effort.

@@

* (Review A) __Some comparison between this paper and [DCH 2003] [DR 2008]?__

  Generally speaking, compared to ML-style modules, families focus on overridability and code inheritance. The latter can be modelled by the functor in a verbose way. Module also has a clear distinction between the implementation and its signature, while family doesn't -- a given family is usually fixed with one signature, closer to OO classes.

  Compared to the paper [DCH 2003], their problem formulation is more mature and they gear towards real-life programming experience. For example, they aim at generativity (a feature for nominality and side effect); subtyping (happens during signature matching); and phase distinction (for compilation). Our current paper handles family in a structural style; doesn't relate different signatures at all; and we work in a full-dependent type setting where mixing static and dynamic phases is acceptable.

  In fact, for plugin development, the family is compiled into modules/functors. In meta-theory, we use sigma type as a conceptually simpler representation of modules. We also use singleton type to expose concrete type information in a family inspired by their work. So our work is heavily influenced by ML-module.

  Our module and mixin have similar semantics to that from [DCH 2003] and [DR 2008]. However, compared to their work, we mainly focus on the prospect of the extensible inductive type and (exhaustiveness checking of) the corresponding recursors. Even in the case of mixin, we consider the consequent mixin of the inductive type and recursors.

### 2023-02-08 pldi2023/review-reply-densed.md

  In fact, for plugin development, the family is compiled into modules/functors. In meta-theory, we use sigma type as a conceptually simpler representation of modules. We also use singleton type to expose concrete type information in a family inspired by their work. So our work is heavily influenced by ML-module.

  Our module and mixin have similar semantics to that from [DCH 2003] and [DR 2008]. However, compared to their work, we mainly focus on the prospect of the extensible inductive type and (exhaustiveness checking of) the corresponding recursors. Even in the case of mixin, we consider the consequent mixin of the inductive type and recursors.

* (Review A/B/C/D) __Section 5 FMLTT is too dense, "incrediably terse" and inappropriately presented to give the audience enough intuition. "More worrisome, the full proof is *only* given in a notation heavy, Agda-style format"__

  Thank you for your feedback. 
  We sympathize with this impression in Section 5, and we agree it is dense, hard to read and not giving too much inspiration to the readers.

  We plan to expand it in the main text to clarify the questions the reviewers have, and also use appendices to explain the formalization and the proof in greater detail aiming for accessibility.

@@

  We will ensure the release of our research artifacts for public access.

* (Review B) __"I think the paper would also be improved by delineating the limitations of the current implementation / metatheory.."__; (Review C) __"Be upfront about this limitation ..."__; (Review D)__ "The biggest of these (which is not mentioned in the paper), is that the presence of open terms complicates typechecking dependently typed terms."__; (Review D) __"It would be nice to add an explicit comment about the trusted code bases of any developments"__

  Thank you for your feedback. We will clearly state the limitations of the
  current implementation and meta-theory in the final revision. 
  <!-- do we want to mention overridability/pins? -->
  In particular, we will emphasize the issue raised by Review D regarding equality coercion
  and the pervasive propositional equality on extensible inductive families.
  This can result in unnecessary equality coercion when `denoteTy t` is not
  definitionally equal to `nat`.

  We believe that this opens up new opportunities for future work in both theory
  and practice, (a). such as a normalization algorithm for an extended calculus
  that admits more conversion equations upon elimination of the extensible
  inductive type; (b). and a proper proof assistant that supports native
  extensible inductive types instead of the current encoding via our plugin,
  thus allowing for proper reduction of open terms. This will change the kernel
  of the proof assistant for sure.

### 2023-02-09 metatheory-condensed.tex
switch citations to author-year format

We contribute \TT, a core type theory that extends
%\underline{f}a\underline{m}i\underline{l}y polymorphism
%\underline{M}artin-\underline{L}öf dependent \underline{t}ype \underline{t}heory (MLTT)~\cite{martin1982constructive}
Martin-Löf dependent type theory (MLTT)~\cite{martin1982constructive}
with facilities to express family polymorphism
while maintaining consistency and canonicity.
%\NA{maybe you can give an overview of the salient points of the design and what they're trying to achieve? What is the core aspect modeled by the theory?}
%
We acknowledge that the presentation herein is dense for an audience
without prior exposure to MLTT, so we summarize the salient points
first before delving into the technical details.

@@

\cref{fig:mltt-selected} presents the syntax and selected typing rules of MLTT
(and \cref{fig:fmltt-selected} \TT).
%We briefly review the base MLTT fragment first.
%
As is the current trend~\cite{coquand2018canonicity,altkap2016,gratzer-sterling-birkedal-2019}
with MLTT presentations, we use \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:
substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}) are part
of the syntax rather than meta-operations.
%
Variables are in the form of \citeauthor{debruijn1964} indices:
\var{n} is the variable bound by the $n$-th closest enclosing binder.
For example, $\lambda x.\,\lambda y.\,x$ is $\lam{\lam{\var{1}}}$.
%
Substitutions are typed with the form 
%\EDJ{It seems to me G |- g : D is a judgement not a rule. Rules usually have two parts (premises + conclusion). Judgment is a standard terminology invented by Martin Lof (https://ncatlab.org/nlab/show/judgment see reference here). I will suggest saying 'Formation rules for substitution lead to a judgement G |- g : D' sth like this. In pen-and-paper, the formation rule (corresponding to QIIT well-typed-ness of judgement) is called presuppose (can be found in Modal MLTT paper) }
$\goodSub{\Gm}{\gm}{\Dl}$.
The idea is that applying~$\gm$ to terms valid in the context~$\Dl$
yields terms valid in~$\Gm$ (\ruleref{tm/sub} and \ruleref{ty/sub}).
%
The two main forms of substitutions are weakening (\ruleref{sub/wk}) and extension (\ruleref{sub/ext}):
\sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
\sub{t}{\SubstExt{\gm}{t'}} substitutes~$t'$ for~$\var{0}$ in~$t$ and
then applies~$\gm$.
For example, rule \ruleref{tm/snd} states that if $t$ is a dependent pair
that has type $\TySigma{A}{B}$, then $\snd{t}$ has type
$\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$,
where $\SubstId$ is the identity substitution (\ruleref{sub/id}).

>>>>

We contribute \TT, a core type theory that extends
%\underline{f}a\underline{m}i\underline{l}y polymorphism
%\underline{M}artin-\underline{L}öf dependent \underline{t}ype \underline{t}heory (MLTT)~\cite{martin1982constructive}
\citeauthor{martin1982constructive} dependent type theory (MLTT)
with facilities to express family polymorphism
while maintaining consistency and canonicity.
%\NA{maybe you can give an overview of the salient points of the design and what they're trying to achieve? What is the core aspect modeled by the theory?}
%
We acknowledge that the presentation herein is dense for an audience
without prior exposure to MLTT, so we summarize the salient points
first before delving into the technical details.

@@

\cref{fig:mltt-selected} presents the syntax and selected typing rules of MLTT
(and \cref{fig:fmltt-selected} \TT).
%We briefly review the base MLTT fragment first.
%
Following the current trend~\cite{coquand2018canonicity,altkap2016,gratzer-sterling-birkedal-2019}
in MLTT presentations, we use \emph{explicit substitutions}~\cite{substcalculus,abadi1989subst}:
substitutions~$\gm$ and their applications (e.g., \sub{T}{\gm}) are part
of the syntax rather than meta-operations.
%
Variables are in the form of \citeauthor{debruijn1964} indices:
\var{n} is the variable bound by the $n$-th closest enclosing binder.
For example, $\lambda x.\,\lambda y.\,x$ is $\lam{\lam{\var{1}}}$.
%
Substitutions are typed with the form 
%\EDJ{It seems to me G |- g : D is a judgement not a rule. Rules usually have two parts (premises + conclusion). Judgment is a standard terminology invented by Martin Lof (https://ncatlab.org/nlab/show/judgment see reference here). I will suggest saying 'Formation rules for substitution lead to a judgement G |- g : D' sth like this. In pen-and-paper, the formation rule (corresponding to QIIT well-typed-ness of judgement) is called presuppose (can be found in Modal MLTT paper) }
$\goodSub{\Gm}{\gm}{\Dl}$.
The idea is that applying~$\gm$ to terms valid in the context~$\Dl$
yields terms valid in~$\Gm$ (\ruleref{tm/sub} and \ruleref{ty/sub}).
%
The two main forms of substitutions are weakening (\ruleref{sub/wk}) and extension (\ruleref{sub/ext}):
\sub{t}{\SubstWeak{n}} introduces $n$ free variables into the context of $t$, and
\sub{t}{\SubstExt{\gm}{t'}} substitutes~$t'$ for~$\var{0}$ in~$t$ and
then applies~$\gm$.
For example, rule \ruleref{tm/snd} states that if $t$ is a dependent pair
that has type $\TySigma{A}{B}$, then $\snd{t}$ has type
$\sub{B}{\SubstExt{\SubstWeak{0}}{\fst{t}}}$,
where $\SubstId$ is the identity substitution (\ruleref{sub/id}).
